---
title: "MCU Box Office Analysis"
output:
  html_document:
      code_folding: hide
      df_print: paged
      
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)

```

```{r, include = FALSE}

# install Packages

#install.packages("rvest")
#install.packages("xml2") # scraping
#install.packages("countrycode")

```

```{r, include = FALSE}

library(rvest)
library(xml2)
library(tidyverse)
library(gridExtra)
library(lubridate)
library(hrbrthemes)
library(png)
library(GGally)
library(patchwork)
library(viridis)
library(ggrepel)
library(scales)
library(countrycode)
library(sf)

```

```{r, include = FALSE}

wiki_mcu <- read_html("https://en.wikipedia.org/wiki/List_of_Marvel_Cinematic_Universe_films")

wiki_mcu %>% 
  html_nodes(css = "table")

```

```{r, include = FALSE}

mcu_table <- 
  wiki_mcu %>% 
  html_nodes(css = "table") %>% 
  nth(6) %>% 
  html_table(fill = TRUE)

head(mcu_table) # Observe first few rows of data

```

```{r, include = FALSE}

# Create data frame from scraped data

mcu <- as.data.frame(mcu_table)


# Remove refs column

mcu <- mcu[,-9]


# Set names of columns

names(mcu) <- c('Film', 'US_release_date', 'Box_office_US_canada', 'Box_office_other', 'Box_office_world', 'Ranking_US_canada', 
                       'Ranking_world', 'Budget')


# Remove unnecessary rows

mcu <- mcu %>%
  filter(!row_number() %in% c(1, 2))


# Change data type of date

date <- as.Date(mcu$US_release_date, "%b %d, %Y")


# Create new phase column

mcu <- mcu %>% 
  mutate(US_release_date = date) %>%
  mutate(Phase =  ifelse (US_release_date < "2013-05-03", "1",
      ifelse (US_release_date >= "2013-05-03" & US_release_date < "2016-05-06", "2",
              ifelse (US_release_date >= "2016-05-06" & US_release_date < "2021-07-09", "3",
                      ifelse (US_release_date >= "2021-07-09", "4", NA)))))


# Remove original "phase" rows/ headers

mcu <- mcu %>%                 
  filter(!grepl('Phase', Film))

    
# Convert box office and budget strings to numeric

mcu <- mcu %>%
  mutate(across(contains("Box"), gsub, pattern = "[^0-9]", replacement = "")) %>% # Remove non numeric symbols 
  mutate(across(contains("Box"), as.numeric)) %>%  # Convert to numeric
  mutate(Budget = gsub("[a-z, $]", "", Budget)) # Remove letters and $ from Budget column


Split <- str_split(mcu$Budget,"–|-") # Note: Have to account for 2 different hyphen characters
Min_Budget <- sapply(Split, "[", 1)
Max_Budget <- sapply(Split, "[", 2)



for(i in 1:length(Max_Budget)){
  ifelse(is.na(Max_Budget[i]) == TRUE, Max_Budget[i] <- Min_Budget[i] , Max_Budget[i] <- Max_Budget[i]) 
} 


mcu <- mcu %>%
  mutate(Min_Budget = as.numeric(gsub("[^0-9.]", "", Min_Budget)) * 1000000) %>% # Convert to numeric 
  mutate(Max_Budget = as.numeric(gsub("[^0-9.]", "", Max_Budget)) * 1000000) %>%
  mutate(Ranking_US_canada = as.numeric(Ranking_US_canada)) %>%
  mutate(Ranking_world = as.numeric(mcu$Ranking_world)) %>%
  select(-Budget)


mcu[nrow(mcu),"Min_Budget"] <- mcu[28,"Min_Budget"]*1000 #Convert total budgets to billions
mcu[nrow(mcu),"Max_Budget"] <- mcu[28,"Max_Budget"]*1000


# Add Year and Month columns

mcu$Year <- year(as.POSIXlt(mcu$US_release_date, format = "%Y-%m-%d"))

months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")

mcu$Month <- months[month(as.POSIXlt(mcu$US_release_date, format = "%Y-%m-%d"))]


# Observe data

glimpse(mcu)

```

```{r, include = FALSE}

# Rotten tomatoes data

site <- "https://editorial.rottentomatoes.com/guide/all-marvel-cinematic-universe-movies-ranked/"
site <- read_html(site)

scores <- html_nodes(site,".tMeterScore") #Used selector gadget via chrome extension
scores <- html_text(scores)


titles <- html_nodes(site,".article_movie_title a")
titles <- html_text(titles)


tomato <- data.frame(Film = titles, tomatometer = scores)



# IMDB data

site2 <- "https://www.imdb.com/list/ls093312123/?sort=list_order,asc&st_dt=&mode=detail&page=1&title_type=movie&ref_=ttls_ref_typ"
site2 <- read_html(site2)

scores2 <- html_nodes(site2,".ipl-rating-star.small .ipl-rating-star__rating") # Used selector gadget via chrome extension
scores2 <- html_text(scores2)

titles2 <- html_nodes(site2,".lister-item-header a")
titles2 <- html_text(titles2)
titles2 <- titles2[1:length(scores2)] # Some films may have no score if they are unreleased


imdb <- data.frame(Film = titles2, IMDB = scores2)

```

```{r, include = FALSE}

# Filter out sony films not contained in the mcu

imdb <- imdb %>% 
  filter(!str_detect(Film, "^Venom")) %>%
  filter(!str_detect(Film, "Morbius")) 


# Rename necessary films to make visualisations more readable and naming schemes consistent

tomato$Film <- as.character(tomato$Film)

tomato$Film[25] <- ("Multiverse of Madness") 
tomato$Film[23] <- ("Age of Ultron")
tomato$Film[21] <- ("Captain America")
tomato$Film[15] <- ("Infinity War")
tomato$Film[12] <- ("The Winter Soldier")
tomato$Film[10] <- ("Far From Home")
tomato$Film[8] <- ("The Avengers")
tomato$Film[6] <- ("Homecoming")
tomato$Film[9] <- ("Shang-Chi")
tomato$Film[5] <- ("No Way Home")
tomato$Film[2] <- ("Endgame")

imdb$Film <- as.character(imdb$Film)

imdb$Film[1] <- ("Endgame")
imdb$Film[2] <- ("Infinity War")
imdb$Film[9] <- ("Captain America")
imdb$Film[10] <- ("The Winter Soldier")
imdb$Film[11] <- ("Age of Ultron")
imdb$Film[14] <- ("Homecoming")
imdb$Film[15] <- ("No Way Home")
imdb$Film[18] <- ("Shang-Chi")
imdb$Film[19] <-("Multiverse of Madness")
imdb$Film[24] <- ("Far From Home")
imdb$Film[25] <- ("Iron Man 3")
imdb$Film[26] <- ("Ant-Man and The Wasp")


# Revert to factor

imdb$Film <- as.factor(imdb$Film)
tomato$Film <- as.factor(tomato$Film)


# Create ratings dataset by merging rotten tomatoes scores and imdb scores

Ratings <- merge(tomato, imdb, by = "Film") 


# Shorten some of the titles in mcu dataset to match with the new names

mcu$Film <- as.character(mcu$Film)


mcu[2,1] <- "The Incredible Hulk"
mcu[5,1] <- "Captain America"
mcu[6,1] <- "The Avengers"
mcu[9,1] <- "The Winter Soldier"
mcu[11,1] <- "Age of Ultron"
mcu[16,1] <- "Homecoming"
mcu[19,1] <- "Infinity War"
mcu[20,1] <- ("Ant-Man and The Wasp")
mcu[22,1] <- "Endgame"
mcu[23,1] <- "Far From Home"
mcu[25,1] <- "Shang-Chi"
mcu[27,1] <- "No Way Home"
mcu[28,1] <- "Multiverse of Madness"

# Revert to factor

mcu$Film <- as.factor(mcu$Film)


```

```{r, include = FALSE}

mcu <- merge(Ratings, mcu, by = "Film") %>%
  arrange(US_release_date)

```
 
```{r, include = FALSE}

Approx_budget <- ((mcu$Max_Budget + mcu$Min_Budget)/2)

mcu$Approx_budget <- Approx_budget

```

```{r, include = FALSE}

mcu$Approx_profit <- mcu$Box_office_world-mcu$Approx_budget

```

```{r, include = FALSE}

mcu <- mcu %>%
  select( - c(Box_office_US_canada, Box_office_other, Ranking_world, Ranking_US_canada ))


```

```{r, include = FALSE}
# Save mcu

saveRDS(mcu, file = "mcu.rds")

mcu <- readRDS(file = "mcu.rds")

mcu

```

```{r, include = FALSE}

ggplot(data = mcu, aes(US_release_date, Box_office_world))+
  geom_point()+
  geom_label_repel(label = mcu$Film, size = 3)+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
  theme_minimal()+ scale_y_continuous(breaks = seq(0, 3000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"))+
  theme_minimal()


```

```{r, include = FALSE}

ggplot(data = mcu, aes(x = reorder(Film, -Box_office_world), y = Box_office_world))+
  geom_bar(stat = "identity", fill = "#00BFC4")+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 3000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"))+
  ylab("Box Office")+
  xlab("US release date")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(axis.title.x=element_blank()) 
  

```

```{r, include = FALSE}

ggplot(data = mcu, aes(x = US_release_date ))+
  geom_line(aes(y= Box_office_world, color = "Box_office_world"))+
  geom_line(aes(y = Max_Budget, color = "Max_Budget"))+
  geom_line(aes(y = Min_Budget, color = "Min_Budget"))+
  scale_colour_manual("", 
                      breaks = c("Box_office_world", "Max_Budget", "Min_Budget"),
                      values = c("black", "red", "orange"))+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
  theme_minimal()+ scale_y_continuous(breaks = seq(0, 3000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"))+
  theme(axis.title.y = element_blank())+
  xlab("US Release Date")
  
  
  
```

```{r, include = FALSE}

  
ggplot(data = mcu, aes(x = reorder(Film, US_release_date), y = Approx_profit))+
  geom_bar(stat = "identity", fill = "#00BFC4")+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 2000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B"))+
  ylab("Estimated profit")+
  xlab("US release date")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x=element_blank()) 



```

```{r, include = FALSE}

# Create a more realistic profit estimate

mcu$Approx_profit_2 <- mcu$Approx_profit - mcu$Approx_budget


# Create a column to check if the release has made or lost money

mcu$In_profit <- ifelse(mcu$Approx_profit_2 > 0, "Yes", "No")

```

```{r, include = FALSE}
  
ggplot(data = mcu, aes(x = reorder(Film, US_release_date), y = Approx_profit_2))+
  geom_bar(stat = "identity", aes(fill = In_profit))+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 2000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B"))+
  ylab("Estimated profit")+
  xlab("US release date")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x=element_blank()) 



```

```{r, include = FALSE}

summary(mcu$Approx_profit_2)

var(mcu$Approx_profit_2)

```

```{r, include = FALSE}

# Read rotten tomatoes image

# Rotten_tomato_image <- readPNG("C:/Users/Adam/Dropbox/Portfolio/MCU_box_office_analysis_files/Rotten_tomatoes.png", native = TRUE)

mcu$tomatometer <- as.numeric(sub("%","",mcu$tomatometer))/100



 ggplot(data = mcu, aes(x = reorder(Film, -as.numeric(tomatometer)), y = tomatometer))+
  geom_bar(stat = "identity", fill = "red")+
      theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(axis.title.x=element_blank())+
   scale_y_continuous(NULL, limits = c(0, 1.0), breaks = seq(0, 1, .2),
                        labels = function(x) scales::percent(x),
                        expand = c(0, 0))+
      geom_text(aes(label = paste0(tomatometer*100,"%")),
                size = 5, 
                hjust = 1.2,
                col = "white",
                angle = 90)
  # inset_element(p = Rotten_tomato_image,   # Combine with image
  #               left = 0.85,
  #               bottom = 0.75,
  #               right = 1,
  #               top = 1)

  

```

```{r, include = FALSE}

 # ggplot(data = mcu, aes(x = reorder(Film, US_release_date), y = tomatometer, group = 1))+
 #  geom_line(color = "#00BFC4")+
 #   expand_limits(y = c(0,1))+
 #   geom_text(aes(label = percent(tomatometer)),
 #                   size = 3, vjust = -0.01) +
 #   scale_y_continuous(NULL, limits = c(0, 1.0), breaks = seq(0, 1, .2),
 #                        labels = function(x) scales::percent(x),
 #                        expand = c(0, 0)) +
 #  theme_minimal()+
 #  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
 #  theme(axis.title.x=element_blank())

```

```{r, include = FALSE}

 # ggplot(data = mcu, aes(x = reorder(Film, US_release_date), y = tomatometer, group = 1))+
 #  geom_line(color = "#00BFC4")+
 #   expand_limits(y = c(0,1))+
 #   scale_y_continuous(NULL, limits = c(0, 1.0), breaks = seq(0, 1, .2),
 #                        labels = function(x) scales::percent(x),
 #                        expand = c(0, 0)) +
 #  geom_label_repel(aes(label = percent(tomatometer)), nudge_x = 0.35, size = 2.5)+
 #  theme_minimal()+
 #  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
 #  theme(axis.title.x=element_blank())#Remove x axis label




```

```{r, include = FALSE}

# Read IMDB image

# IMDB_logo <- readPNG("C:/Users/Adam/Dropbox/Portfolio/MCU_box_office_analysis_files/IMDB_logo.png", native = TRUE)


mcu$IMDB <- as.double(as.character(mcu$IMDB))




ggplot(data = mcu, aes(x = reorder(Film, -IMDB), y = IMDB))+
  geom_bar(stat = "identity", fill = "orange")+
  expand_limits(y = c(0,10))+
  ylab("Score")+
  geom_text(aes(label = IMDB), 
            hjust = 1.5, 
            colour = "white",
            size = 5,
            angle = 90)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.title.x=element_blank())
  # inset_element(p = IMDB_logo,   # Combine with image
  #               left = 0.6,
  #               bottom = 0.7,
  #               right = 1.1,
  #               top = 1.1)

```

```{r, include = FALSE}


ggplot(mcu, aes(x = reorder(Film, US_release_date))) +
  geom_line( aes(y = tomatometer, group = 1), color = "Red") + 
  geom_line( aes(y = IMDB / 10, group = 1), color = "orange") + 
  scale_y_continuous(name = "Tomatometer", labels = function(x) scales::percent(x), sec.axis = sec_axis(~.*10, name="IMDB score"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
        ,axis.title.y = element_text(color = "red", size=13, vjust = 3)
        ,axis.title.y.right = element_text(color = "orange", size=13, vjust = 3))+
  theme(axis.title.x=element_blank())
  
```

```{r, include = FALSE}

summary(mcu$tomatometer)

```

```{r, include = FALSE}

summary(mcu$IMDB)

```

```{r, include = FALSE}

ggplot(mcu, aes(x = reorder(Film, US_release_date))) +
  
    geom_col( aes(y = Box_office_world, group = 1), fill = "#00BFC4")+
  geom_line( aes(y = tomatometer*2000000000 , group = 1), color = "Red") +
  
  scale_y_continuous( name = "Box office", breaks = seq(0, 3000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"),
                     sec.axis = sec_axis(~./2000000000, name = "Tomatometer", labels = function(x) scales::percent(x)))+
  
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
        ,axis.title.y = element_text(color = "#00BFC4", size=13)
        ,axis.title.y.right = element_text(color = "red", size=13, vjust = 3))+
  theme(axis.title.x=element_blank())


```

```{r, include = FALSE}

cor(mcu$Box_office_world, mcu$tomatometer)

```

```{r, include = FALSE}

ggplot(mcu, aes(x = reorder(Film, US_release_date))) +
  
    geom_col( aes(y = Box_office_world, group = 1), fill = "#00BFC4")+
  geom_line( aes(y = IMDB*200000000 , group = 1), color = "orange") +
  
  scale_y_continuous( name = "Box office", breaks = seq(0, 3000000000, by = 500000000), 
                     labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"),
                     sec.axis = sec_axis(~./200000000, name = "IMDB"))+
  
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
        ,axis.title.y = element_text(color = "#00BFC4", size=13)
        ,axis.title.y.right = element_text(color = "orange", size=13))+
  theme(axis.title.x=element_blank())

```

```{r, include = FALSE}

cor(mcu$Box_office_world, mcu$IMDB)

```

```{r, include = FALSE}

phase <- data.frame(start = as.Date(c("2008-05-02", "2013-05-03", "2016-05-06", "2021-07-09"), format =  "%Y-%m-%d"),  
                    end = as.Date(c("2013-05-03", "2016-05-06", "2021-07-09", "2023-02-01"), format = "%Y-%m-%d"),
                    Phase = factor(1:4))

phase


```

```{r, include = FALSE}
mcu$cumu_profits <- cumsum(mcu$Approx_profit_2)

```

```{r, include = FALSE}

ggplot() + 
  geom_rect(data = phase,  aes(xmin=start, xmax = end, 
                 ymin = -Inf, ymax = Inf, 
                 fill = Phase), alpha = 0.5)+
  scale_fill_viridis(discrete = TRUE, option = "C")+
  geom_point(data = mcu, aes(US_release_date, cumu_profits))+
  geom_line(data = mcu, aes(US_release_date, cumu_profits))+
  scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-10, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 20000000000, by = 5000000000), 
                     labels = c(0,"5 B", "10 B", "15 B", "20 B"))+
  ylab("Cummulative profits")+
  xlab("US release date")+
  theme_minimal()
  
```

```{r, include = FALSE}

mcu_year <- mcu %>% 
              group_by(Year) %>%
              summarise(Mean_box_office = mean(Box_office_world))

mcu_month <- mcu %>% 
              group_by(Month) %>%
              summarise(Mean_box_office = mean(Box_office_world))


mcu_year; mcu_month


```

```{r, include = FALSE}

Releases_per_year <- ggplot(data = (mcu %>%count(Year)), aes(x = Year, y = n) )+
                      geom_bar(stat = "identity", fill = "#00BFC4")+
                        scale_x_continuous(breaks=2008:2022)+
                        theme_minimal()+
                        ggtitle("Releases per year")+
                        theme(plot.title = element_text(hjust = 0.5),
                        axis.title.y = element_blank(),
                        axis.text.y = element_blank(),
                        axis.title.x = element_blank(),
                        plot.margin = unit(c(1,-1,1,0), "mm"))+
                        scale_y_reverse()+
                        coord_flip()


Mean_box_ofice_year <- ggplot(data = mcu_year, aes(x = Year, y = Mean_box_office))+
                        geom_bar(stat = "identity", fill = "#00BFC4")+                        
                        scale_x_continuous(breaks=2008:2022)+
                        scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
                        scale_y_continuous(breaks = seq(0, 3000000000, by = 500000000), 
                                           labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"))+
                        ylab("Box Office")+
                        theme_minimal()+
                        ggtitle("Mean box ofice per year") +
                        theme(plot.title = element_text(hjust = 0.5),
                        axis.title.x = element_blank(),
                        axis.title.y = element_blank(),
                        plot.margin = unit(c(1,-1,1,0), "mm"))+
                        coord_flip()


grid.arrange(Releases_per_year, Mean_box_ofice_year, ncol = 2)

```

```{r, include = FALSE}

month_count <- mcu %>% count(Month)
               

month_count

# Create abreviated month data frame 

month_abv <- data.frame(Month = c("Jan","Feb","Mar","Apr","May","Jun", "Jul","Aug","Sep","Oct", "Nov", "Dec"))


# Add missing months 

month_modified <- left_join(month_abv, month_count, by = "Month") 

 
# Replace coerced NAs with 0 

month_modified <- month_modified %>%
  mutate(n = ifelse(is.na(n),0,n)) %>%
  mutate(Month = factor(Month, levels = c("Jan","Feb","Mar","Apr","May","Jun", "Jul","Aug","Sep","Oct", "Nov", "Dec")))


# Create plot

month_plot <- ggplot(month_modified, aes(x = Month, y = n)) +
                geom_bar(stat = "identity", fill = "#00BFC4")+
                scale_y_reverse()+
                ggtitle("Releases per month")+
                theme_minimal()+
                theme(plot.title = element_text(hjust = 0.5),
                      axis.title.y = element_blank(),
                      axis.text.y = element_blank(),
                      axis.title.x = element_blank())+
                coord_flip() +
                scale_y_reverse()


```

```{r, include = FALSE}

# Add missing months 

month_modified_mean <- left_join(month_abv, mcu_month, by = "Month") 

# Replace coerced NAs with 0 

month_modified_mean <- month_modified_mean %>%
  mutate(Mean_box_office = ifelse(is.na(Mean_box_office),0,Mean_box_office)) %>%
  mutate(Month = factor(Month, levels = c("Jan","Feb","Mar","Apr","May","Jun", "Jul","Aug","Sep","Oct", "Nov", "Dec")))


# Create plot

month_mean_plot <- ggplot(month_modified_mean, aes(x = Month, y = Mean_box_office)) +
                geom_bar(stat = "identity", fill = "#00BFC4")+
                ggtitle("Mean box office")+
                        scale_y_continuous(labels = unit_format(unit = "B", scale = 1e-9, accuracy = 0.5))+
                        scale_y_continuous(breaks = seq(0, 3000000000, by = 500000000), 
                                           labels = c(0,"500 M", "1 B", "1.5 B", "2 B", "2.5 B", "3 B"))+
                theme_minimal()+
                theme(plot.title = element_text(hjust = 0.5),
                      axis.title.y = element_blank(),
                      axis.title.x = element_blank())+
                coord_flip()


            

```



## Part 3 - Box office of Endgame by country

$$\\[0.2in]$$

In part 3, we will examine the box office returns of the highest grossing MCU film "Avengers: Endgame" (2019) as it relates to the countries it was released in. The data we use here is available on boxofficemojo.com and contains the film's release date, the market (i.e. the country it was released in), the earning over the opening weekend the gross earnings at the box office.

Firstly, the data is collected and cleaned.


```{r class.source = 'fold-show',eval = FALSE}

# Scrape multiple tables 

library(rvest)

url <- "https://www.box officemojo.com/releasegroup/gr3511898629/"

endgame <- url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table(fill = TRUE)


# Convert list of dataframes into a single dataframe

endgame <- bind_rows(endgame)


endgame[3:nrow(endgame), 1:16] <- endgame[3:nrow(endgame), 5:ncol(endgame)]

endgame[36:nrow(endgame), 1:16] <- endgame[36:nrow(endgame), 5:ncol(endgame)]

endgame[46:nrow(endgame), 1:16] <- endgame[46:nrow(endgame), 5:ncol(endgame)]

endgame[61:nrow(endgame), 1:16] <- endgame[61:nrow(endgame), 5:ncol(endgame)]

endgame[,5:ncol(endgame)] <- NULL


# Rename column names

names(endgame) <- c('Market', 'Release Date', 'Opening', 'Gross')


# Clean up dataframe

endgame <- endgame %>%
  filter(!Market %in% c("Market") 
         & !`Release Date` %in% c("Release Date"))     # Note: `` allows indexing column with a space in name


# Convert box office to numeric

endgame$Gross <- as.numeric(gsub('[$,]', '', endgame$Gross))
endgame$Opening <- as.numeric(gsub('[$,]', '', endgame$Opening))


# Change some country names to match with map data 

library(sf)
library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

endgame$Market[which(!(endgame$Market %in% world$name))]

endgame$Market[1] <- "United States"
endgame$Market[6] <- "Czech Rep." 
endgame$Market[23] <- "Serbia"
endgame$Market[30] <- "Turkey"
endgame$Market[51] <- "Russia"
endgame$Market[53] <- "Korea"


# Add regions

library(countrycode)

endgame$Region <- countrycode(sourcevar = as.factor(endgame$Market), origin = "country.name", destination = "region")


# Rename United States to include Canada

endgame$Market[1] <- "United States & Canada"


# Remove release date from data frame

endgame <- endgame %>%
  select(-"Release Date")

```

Save and observe data

```{r}

# Save endgame
#saveRDS(endgame, file = "endgame.rds")

endgame <- readRDS(file = "endgame.rds")

endgame

```

$$\\[0.5in]$$

We will take an initial look at the gross earnings over various regions.


```{r}

ggplot(endgame, aes(x = reorder(Region, Gross), y = Gross))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = NULL)+
  coord_flip()+ 
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-8, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 1000000000, by = 200000000), 
                     labels = c(0,"200 M", "400 M", "600 M", "800 M", "1 B"))+
  theme_minimal()

```

$$\\[0.2in]$$

Clearly the regions "North America" and "East Asia & Pacific" account for a large amount of the earnings. "North America" only has one data point as the box office of Canada is included in the domestic box office for North America. In the case of East Asia, we observe an outlier bringing in over $600 million, considerably more than the other countries in the region. The Middle East, North Africa and Sub-Saharan Africa appear to have generated the least money for the film.

$$\\[0.5in]$$

Let's examine which specific countries have taken in the most money for the film. For example, the top 20 earners.

```{r}

#Top 20 box offices

# Create magma colour palette

mag_pal <- viridis(n = 7)

cols <- c("North America" = mag_pal[1], "East Asia & Pacific" = mag_pal[2], "Europe & Central Asia" = mag_pal[3], "Latin America & Caribbean" = mag_pal[4], "South Asia" = mag_pal[5], "Middle East & North Africa" = mag_pal[6], "Sub-Saharan Africa" = mag_pal[7])


# Select top 20 performing countries

endgame_1 <- arrange(endgame, -Gross)[1:20,]

endgame_1

```

$$\\[0.2in]$$

We will also calculate the combined earnings of the top two performing countires as a percentage of the whole: 

```{r}

# % of US and China together

(sum(endgame_1$Gross[1:2])/sum(endgame_1$Gross)) * 100


```



```{r}

ggplot(endgame_1, aes(x = reorder(Market, Gross), y = Gross, fill = Region))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-8, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 1000000000, by = 200000000), 
                     labels = c(0,"200 M", "400 M", "600 M", "800 M", "1 B"))+
  scale_fill_manual(values = cols)+
  theme_minimal()+
  labs(x = NULL)+
  coord_flip()+
  ggtitle("Top 20 countries")

```
$$\\[0.5in]$$

Here we note two standout countries where the film performed the best, namely the United States where it brought in over $850 million and China with just under $630 million. These two countries alone are responsible for 60.43% of Endgame's box_office earnings.

They are followed by the UK ($115 million), Korea($105 million) and Brazil ($85 million).

Countires in the regions of "Middle east & North Africa" and "Sub-Saharan Africa" are not included in the list of top 20 earners.


$$\\[0.5in]$$

When we divide the Gross earnings by region we observe that East Asia and the pacific take in the most money, followed by North America as expected.

```{r}

# Group data by region

endgame_3 <- endgame %>%
  group_by(Region) %>%
  summarise(Gross = sum(Gross)) %>%
  arrange(-Gross)

# Observe data

endgame_3

```

$$\\[0.5in]$$

```{r}

# Visualise this data

# Calculate polar coordinates for piechart labels

df <- endgame_3 %>% 
  arrange(Gross) %>%
  mutate(csum = rev(cumsum(rev(Gross))), 
         pos = Gross/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Gross/2, pos))


# Create plot

ggplot(endgame_3, aes(x="", y = Gross, fill = reorder(Region, Gross))) + #Reorder - piechart goes from highest value to lowest clockwise
  geom_bar(stat="identity", width = 1, color = "white")+
  ggtitle("Gross by Region")+
  coord_polar("y", start=0) +
  geom_label_repel(data = df,
                   aes(y = pos, label = paste0(round(Gross / sum(Gross) * 100, 1), "%")),
                   size = 5, nudge_x = 1, show.legend = FALSE, color = "darkorange") +
  scale_fill_manual(values = cols)+
  theme_void()+
  theme(plot.title = element_text(hjust = 0.5, vjust = 3)) +
  guides(fill = guide_legend(title = "Region")) 


```

$$\\[0.5in]$$

Taking a closer look at East Asia and the Pacific, unsurprisingly it seems that it is China that accounts for the majority of box office earnings.

```{r}

endgame_Asia <- endgame %>% 
                filter(Region == "East Asia & Pacific")%>%
                arrange(-Gross) %>%
                select(Market, Gross) 

endgame_Asia

ggplot(endgame_Asia, aes(reorder(Market, -Gross), Gross))+
  geom_bar(stat = "identity")+
  geom_col(fill = "#00BFC4")+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-8, accuracy = 0.5))+
  scale_y_continuous(breaks = seq(0, 1000000000, by = 200000000), 
                     labels = c(0,"200 M", "400 M", "600 M", "800 M", "1 B"))+
  ggtitle("East Asia & Pacific")


```


$$\\[0.5in]$$

To get a better sense of the geographical data available to us, we can visualise the endgame dataset on a world map. Here countries with available data are highlighted. Once again, note that the Canadian box office is contained in the domestic box office for North America.


```{r class.source = 'fold-show', paged.print=FALSE}

# Generate simple feature collection for visualisation

library("sf")
library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

# Combine geometry of Canada and USA to create "North America"

North_America <- world %>%
 filter(name == "Canada" | name == "United States") %>%
  st_combine()

data <- data.frame(name = "United States & Canada", subregion = "North America")

North_America_2 <- st_sf(data, geometry = North_America)


# Observe data

North_America_2


sfc <- world %>%
  select(name, geometry, subregion) 

sfc <- rbind(sfc, North_America_2)


# Observe data 

sfc


world_joined <- sfc %>%
  left_join(endgame, by = c("name" = "Market"))%>%
  filter(name != "Antarctica")

```

```{r}

# Create heat map of profits

ggplot() + 
 geom_sf(data = world_joined, aes(fill = Gross))+
 ggtitle("Gross by country") +
 theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())+
 scale_fill_continuous(name = "Gross", breaks = seq(0, 1000000000, by = 250000000), 
 labels = c(0,"250 M", "500 M", "750 M", "1 B"))

```

Generating a heat map based on box office performance again highlights the dominance of North America and China, it also allows us to see at a glance the countries that were included and excluded from the data.


$$\\[0.5in]$$

Our data  not only contains the gross box office by country but also the money made over the film's opening weekend. Studios and critics will often emphasize the importance of a strong opening weekend as it is a good indication of how well the movie will perform over its run.

We will visualise how the opening performance of the film compares to it's overall box office. 

```{r}

# Create dataset using pivot_longer

endgame_2 <- endgame_1 %>%
  select(Market, Opening, Gross) %>%
  mutate("Remaining" = Gross - Opening ) %>%
  pivot_longer(cols = c("Opening", "Remaining"), names_to = "Box_Office", values_to = "Amount") %>%
  na.omit() # Revove Hong Kong and Korea where opening is unavailable
  
 
# Observe data

 endgame_2


# Create plot
 
ggplot(endgame_2, aes(x = reorder(Market, Gross), y = Amount, fill = forcats::fct_rev(Box_Office)))+ #forcats::fct_rev places "Opening" at the left 
  geom_bar(position = "fill", stat = "identity")+
  ggtitle("Opening & Remaining Box Office")+
  scale_y_continuous(labels = scales::percent)+
  theme_minimal()+
  labs(x = NULL)+
  geom_text(aes(label = paste0(round((Amount/Gross)*100,2),"%")),
                size = 4, 
                position = position_fill(vjust = 0.4), 
                col = "white")+
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.title = element_blank())



```

$$\\[0.2in]$$

Interestingly, a significant proportion of the overall box office appears to come from the opening weekend alone.

In the case of the United States, 41.6% of the earnings were obtained in this short time period. In the UK, the third highest performing country, Endgame brought in just under 49% of it's gross over the opening. 

The film was screened in cinemas in North America from April 26th to September 12th (140 days), meaning it brought in 41.6% of it's money only 2.14% (three days) of the way into it's total run.

Even the lowest opening percentage - Japan, still accounts for almost a quarter(23.83%) of the total box office. As such, this plot helps underline the importance of a solid opening for a blockbuster movie. 

Note that Korea and Hong Kong have been omitted here as the data for their opening weekend box office was unavailable




$$\\[1.5in]$$






